%!TEX root = ../thesis.tex
\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
%
This work constitutes my master thesis on how well different machine learning models can improve on a mathematical programming algorithm. The work is a part of a master's degree in \textit{Engineering Cybernetics}, with the specialization\textit{ Autonomous System Control}.
The work equates to 30 ECTS-credits, equal to one semester. 

I have completed the experiments using only free and open software, and all experiments were run on computer hardware I have paid for myself.  

The thesis is based on the work from my specialization project, \textit{Multi-Layer Perceptrons for Learning to Branch
} (2020), and chapters, sections, and figures are taken or adapted from that report. These sections will be indicated at the beginning of each chapter, but some minor adaptions will not be commented on in order to keep the thesis unencumbered by excessive comments.

The idea for the thesis as well as the research questions was formulated by myself.
While not officially affiliated with the \gls{DS4DM} group at the Polytechnique Montr\'{e}al, their continuous work on the \gls{Ecole} framework has been instrumental to the thesis. My questions and suggestions were answered immediately, and their cooperation was very important for the success of the project. I would like to specifically mention Maxime Gasse, Didier Ch\'{e}lat, and Antoine Provoust for their help, as well as other researchers and programmers who make their work publicly available. I hope to make a contribution in that regard by also making the code for this thesis open on GitHub.

Lastly, thank you to my supervisor professor Lars Struen Imsland and my co-supervisor Bjarne Grimstad, with whom I have had discussions regarding the thesis. 


%
\clearpage
